\documentclass[a4paper,11pt]{esiwace-modified}

%\input{esiwace_packages}

\title{D4.2 New Storage Layout for Earth System Data}
\author{Jakob Lüttgau
	\and Julian Kunkel
	\and Bryan Lawrence
	\and Alessandro D'Anca
	\and Paola Nassisi
	%\and Jens Jensen
	\and Giuseppe Congiu
	\and Huang Hua
	\and Sandro Fiore
	\and Neil Massey
}

\date{\today} % Replace with final date

\ESIWACEworkpackage{WP4 Exploitability}
\ESIWACElead{DKRZ}
\ESIWACEparticipants{Seagate, CMCC, STFC, UREAD}




%\usepackage[inline]{showlabels}

\usepackage{float}

\graphicspath{
    {../figures/}
    {../../../template/images/}
    {../images/}
    {../assets/}
    {../../assets/}
}

\usepackage{adjustbox}
\usepackage{enumitem}

\setlength{\parindent}{0pt}



\def\myversion{0.2.5}
\def\approvaldate{2017-06-31}
%
% %%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%
%
\begin{document}
\title{ESD Middleware Architecture}
\date{\today}

\maketitle

\tableofcontents

\begin{abstract}
Making the best use of HPC in Earth simulation requires storing and manipulating vast quantities of data.
Existing storage environments face usability and performance challenges for both domain scientists and the data centers supporting the scientists.

These challenges arise from data discovery/access patterns, and the need to support complex legacy interfaces. In the ESiWACE project, we develop a novel I/O middleware targeting, but not limited to, earth system data.
This deliverable sheds light upon the technical design of the ESD middleware, and the user perspective and implications when using the middleware.
Its architecture builds on well-established end-user interfaces but utilizes scientific metadata to harness a data structure centric perspective.

In contrast to existing solutions, the middleware maps data structures to available storage technology based on several parameters:
1) A data center specific configuration of available hardware with their characteristics;
2) The intended usage pattern explicitly provided by the user and implicitly by the structure of the data.

This allows exploiting performance characteristics of a heterogeneous storage environment more efficiently.

This deliverable provides the background on data representations and description formats commonly used in earth system modeling.
The document isolates the key requirements for an earth system middleware and collects numerous use-case outlining the benefit to existing and anticipated workflows and technologies.
Finally, a detailed initial design for the architecture of the earth system middleware is proposed and documented.

The document is not intended to describe all components completely but provides a high-level overview that is necessary to build a first prototype as it is planned in the next phase of the ESiWACE project.
During this development, the design will be adjusted to match the prototype; the final version of the design document will be delivered with the end of the project.
\end{abstract}



\chapter*{Revision History}

\begin{center}
\begin{tabular} {| c | l | l | l | } \hline
    Version & Date & Who & What  \\ \hline
    0.2.5 & July 3rd, 2017 & Team & Architecture draft. \\
    0.3 & December 18th, 2018 & Team & Correction of various typos and minor flaws. \\
    \hline
\end{tabular}
\end{center}


\chapter{Introduction}
\begin{chapterIntro}
This document provides the architecture for our new Earth System Data Middleware (\textbf{ESDM})\footnote{Depending on the context, we may use as full name ESD middleware.},
aimed at deployment in both simulation and analysis workflows where the volume and rate of data leads to performance and data management issues with traditional approaches.
This architecture is one of the deliverables from the Centre of Excellence in Weather and Climate in Europe (http://esiwace.eu).
\end{chapterIntro}


\section{General Objectives}

In this section we outline the general challenges, and some specific challenges which this work needs to address.
Detailed consequential requirements appear in \Cref{sec:requirements}.

\subsection{Challenges and Goals}
\label{sec:esd goals}
\input{intro_challenges.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Architecture Philosophy and Methodology}
\label{sec:about 4+1}
\input{intro_methodology.tex}

\section{Document Structure}

Before delving into the formal software architecture from a software engineering perspective,
we introduce some key aspects of background information about data layout and data formats which provide context for both actual architectural decisions and some directions in which the architecture might evolve.
\Cref{sec:background} concludes with a description of key storage components which we consider for targeting in the architecture proper.

We extract the general properties of requirements from the Logical View and present them in \Cref{sec:requirements}, where we also introduce elements
of related work which a priori influence the architecture itself (e.g.to introduce why we have introduced specific third party dependencies).
\Cref{chap:use cases} isolates use cases for the ESDM which, in turn, drive the architecture discussion.
\Cref{chap:viewpoints} proceeds with the architecture properties, beginning with an overview, before addressing the various viewpoints.
\Cref{chap:components and backends} addresses the scenarios and use cases, including the first implementation scenarios that will be necessary to meet ESIWACE requirements.

The document concludes with a summary chapter (\Cref{chap:summary}) which relates the specific functional requirements to specific aspects of the architecture.



 \chapter{Background}
\label{sec:background}
\input{background_formats.tex}

\input{background_detailed.tex}




%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Requirements}
\label{sec:requirements}
\input{requirements.tex}

%\section{Related Work and Dependencies}
%In this section we discuss the influence of important third party architectures on the ESD requirements.
%\label{sec:related}
% In which we discuss relationships to key projects such as JR's HDF server activity, ophidia etc.
%\input{requirements_related.tex}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Use-Cases}
\label{chap:use cases}
\input{use-cases}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Architecture: Viewpoints}
\label{chap:viewpoints}
\input{viewpoints.tex}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Architecture: Components and Backends}
\label{chap:components and backends}
\input{backend.tex}




%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Summary}
\label{chap:summary}

The primary goal of this deliverable is to provide an initial architecture design for the ESD middleware.

It begins with some introductory material on the scientific context, and the dominant APIs and file formats (chapter \ref{sec:background}).
The applications tend to use structured grids and common file formats are NetCDF, HDF5 and GRIB.
Most applications are parallelized using MPI which sometimes also influences the data model of applications.
Parallel file systems are the most common storage system at the moment but object storage is becoming more and more important.
New technologies predominantly used in ``big data'' contexts may have benefits for scientific data storage and workflows.

The architecture design exploits the 4+1 architecture method: 4 different architectural views accompanied by use cases as drivers.
The scene for the design is set by a chapter on requirements (chapter \ref{chap:use cases}), which lists both functional and non-functional requirements on top of the simple need for CRUD (Create, Retrieve, Update, Delete) for both data and metadata.
The architecture has to support heterogeneous architectures and should be compatible to existing technologies.
In addition, the resulting software should be configurable, user-friendly, reliable and performant.

In chapter \ref{chap:use cases}) use cases covering the most common tasks in existing climate and weather workflows are collected and
the different stakeholders (e.g. scientists, project managers, institutions which provide infrastructure) are described.
The core systems such as the supercomputers, storage systems, applications, and software libraries are briefly introduced and characterized by the main failure modes to be considered by the middleware.
The body of the chapter comprises a list of high-level basic use cases which are analysed using a consistent approach to identify key characteristics.
In each case a flow of events which includes the interaction with the Earth System Data Middleware is identified.
The cases covered include conventional simulation runs, independent pre/post-processing jobs as well as more integrated scenarios that implement pipelines, workflows and in situ methods.

A general architecture overview in chapter \ref{chap:viewpoints} is driven by the challenges and requirements collected earlier, and provides a high level view of the first three of the formal architectural views (logical, physical, process).
The logical view covers the components, data model, expected operations and semantics.
The physical view covers the relationship between the software components and the hardware components, while the process view describes the active components, their interaction, and how they drive the I/O.
Key aspects of the conceptual data model are the description of multi-dimensional data sets, variables and their metadata.
Alongside typical operations to access and manipulate data asynchronously, the architecture covers the necessary notifications needed to enable management communication between the middleware and the system.

\Cref{chap:components and backends} refines the high level views introduced in chapter \ref{chap:viewpoints}, providing more detail.
A scheduling component is responsible to accept and delegate incoming requests to the layout component as well as the different backends.
The layout component finds suitable data fragmentation and bases this decision on a performance model which also reflects the different storage backends.
The storage backends in turn are responsible to store fragments but also to provide a performance estimate that matches the storage system and its configuration.
Multiple storage backends are discussed including POSIX-like systems as well as two solutions for object storage.

The document is not intended to describe all components completely but provides a high-level overview that is necessary to build a first prototype; as such there is no Development View included here.
The development view will come with the prototype, and it is expected that the architecture design will then iterate with actual development and deployment experience.





%\section{Systems Deployments}
%\action{BNL\&JK:List of scenarios}{BNL to discuss with JK an initial list of deployment scenarios}
%\begin{Discussion}
%	ESD Demonstrator. GridFTP direct to ESD interface. Which one?
%	Where deployed. Where is the ESD storage. Cache. Object Store etc.
%\end{Discussion}




%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%\bibliographystyle{alpha}
\bibliography{bibliography.bib}
\label{sec:references}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{use-cases-appendix}


\end{document}
