
\begin{chapterIntro}
	This section discusses most relevant components and backends in technical detail.
	For every component listed, the 4+1 view is described (refer to \Cref{sec:about 4+1}).
	The components and backends are ordered according to their position within the I/O stack from top (application) to bottom (backend).

	\Cref{component: scheduler} discusses the scheduling component that breaks down incoming read and write requests into subsequent requests to create or receive multiple fragments.
	The scheduler relies on a layout component to create a set of domain filling subrequests which is discussed in \Cref{component: layout}.
	In most cases, an application will not interface directly with the ESDM middleware but through a common frontend.
	\Cref{frontend: hdf5 + mpi} introduces a HDF5 frontend and also discusses how message passing via MPI is realized outside of the ESDM.
	\Cref{frontend: fuse} addresses legacy interfaces using FUSE to expose datasets via a configurable virtual file systems.

	It follows the discussion of multiple backends.
	In particular \Cref{backend: posix} discusses a POSIX backend to allow for interactions with parallel file systems.
	Object storage backends for Mero (see \Cref{backend: mero}) and WOS (see \Cref{backend: wos}) are discussed,
	Besides data backends, also pure metadata backends are possible allowing to use a existing software stacks that are typically also residing on some kind of storage backend themself.
	\Cref{backend: mongo} describes MongoDB bases metadata backend.
\end{chapterIntro}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Scheduling Component}
\label{component: scheduler}
\input{component-scheduler}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage
\section{Layout Component}
\label{component: layout}
\input{component-layout}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage
\section{HDF5+MPI plugin}
\label{frontend: hdf5 + mpi}
\input{frontend-hdf5+mpi}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage
\section{Fuse Legacy + Metadata Mapped Views}
\label{frontend: fuse}
\input{frontend-fuse}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage
\section{Backend POSIX/Lustre (Using ESDM)}
\label{backend: posix}
\input{backend-posix}


\section{Mongo DB Metadata backend}
\label{backend: mongo}
\input{backend-metadata-mongo.tex}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage
\section{Mero Backend}
\label{backend: mero}
\input{backend-mero}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage
\section{WOS Backend}
\label{backend: wos}
\input{backend-wos}
